Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}

### 2008-04-17 docs/Godunov_notes/paper.tex
Merged the rest of the enthalpy_pred_type tables into the document

@@

Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
u_{R,\ib-\half\eb_x}^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
\end{eqnarray}
Then, solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0.
\end{cases}
\end{equation}
Use the exact analogous procedure to obtain $v^{\mac}_{\ib-\half\eb_y}$.
%\subsection{3D Cartesian Case}
\cleardoublepage
\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.  To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.  
\subsection{2D Cartesian Case}
The 1D extrapolation is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \\
\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\mac}$:
\begin{equation}
\rho_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - \rho_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
&& \\
\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - \rho_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
\end{eqnarray}
Then, we upwind based on $u^{\mac}$.
\begin{equation}
\rho_{\ib-\half\eb_x}^{\edge} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
\end{cases}
\end{equation}
%\subsection{3D Cartesian Case}
\cleardoublepage
\section{Notes on {\tt enthalpy\_pred\_type} in {\tt MAESTRO}}

\begin{table*}
\begin{center}
\caption{When predicting temp ... \newline}
\begin{tabular}{|c|c|c|c|c|}
\multicolumn{1}{c}{PRED X} & \multicolumn{1}{c}{type} & \multicolumn{1}{c}{Inputs } & 
\multicolumn{1}{c}{EOS inputs} & \multicolumn{1}{c}{output} \\
\hline
T & 3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
  &   &                   & $   X = $ given & \\
  &   &                   &                 & \\
T & 4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
  &   &                   & $   X = $ given & \\
  &   &                   &                 & \\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\caption{Quantity that goes into mkflux on edges \newline}
\begin{tabular}{|c|c|c|c|c|}
\multicolumn{2}{c}{PREDICT} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
\multicolumn{1}{c}{X} & \multicolumn{1}{c}{h pred type} & \multicolumn{1}{c}{Species } & \multicolumn{1}{c}{h/T } & \multicolumn{1}{c}{$\rho^\prime$} \\
\hline
T & 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ & \\
T & 2 & $X$ and $\rho^\prime$ & $h$ & \\
T & 3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ & \\
T & 4 & $X$ and $\rho^\prime$ & $h$ & \\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\caption{Quantity that is created in mkflux and passed into update scal\newline}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\multicolumn{2}{c}{PREDICT}   & 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{}  
& \multicolumn{2}{c}{Add db?} & \multicolumn{2}{c}{Forcing term }\\
\multicolumn{1}{c}{X}   & \multicolumn{1}{c}{type} &
\multicolumn{1}{c}{Species } & \multicolumn{1}{c}{$(\rho h)$} 
& \multicolumn{1}{c}{X} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{$(\rho X)$} & \multicolumn{1}{c}{$(\rho h)$} \\
\hline
T & 1 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)(\rho h)^\prime + \ut (\rho h)_0$ & N & Y & 0 & $ (\tilde{w} \frac{\partial p_0}{\partial r}) $ \\
T & 2 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & N & N & 0 & $ (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) $ \\
T & 3 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)(\rho h)^\prime + \ut (\rho h)_0$ & N & Y & 0 & $ (\tilde{w} \frac{\partial p_0}{\partial r}) $ \\
T & 4 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & N & N & 0 & $ (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) $ \\
\hline
\end{tabular}
\end{center}
\end{table*}

\end{document}

>>>>

If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}$ to $h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.  Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.

@@

Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
\begin{eqnarray}
u_{L,\ib-\half\eb_x}^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
u_{R,\ib-\half\eb_x}^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
&-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
\end{eqnarray}
Then, solve a Riemann problem:
\begin{equation}
u_{\ib-\half\eb_x}^{\mac} =
\begin{cases}
0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0.
\end{cases}
\end{equation}
Use the exact analogous procedure to obtain $v^{\mac}_{\ib-\half\eb_y}$.
%\subsection{3D Cartesian Case}
\cleardoublepage
\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.  To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.  
\subsection{2D Cartesian Case}
The 1D extrapolation is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \\
\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n.
\end{eqnarray}
Then we upwind based on $u^{\mac}$:
\begin{equation}
\rho_{\ib-\half\eb_x}^{1D} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
\end{cases}
\end{equation}
Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
\begin{eqnarray}
\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - \rho_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
&& \\
\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - \rho_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
\end{eqnarray}
Then, we upwind based on $u^{\mac}$.
\begin{equation}
\rho_{\ib-\half\eb_x}^{\edge} =
\begin{cases}
\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
\end{cases}
\end{equation}
%\subsection{3D Cartesian Case}
\end{document}
